---
title: "A3_MachineLearning"
author: "Jakob, Emilie, Gergana & Natasha"
date: "2022-11-02"
output: html_document
---

# The assignment

The Machine Learning assignment has 3 main parts: 
- First we create a skeptical and an informed simulation, based on the meta-analysis. 
- Second we build and test our machine learning pipeline on the simulated data. 
- Second we apply the pipeline to the empirical data.

The report for the exam, thus, consists of the answer to all the following prompts:
- Describe your machine learning pipeline. 
- Produce a diagram of it to guide the reader (e.g. see Rybner et al 2022 Vocal markers of autism: Assessing the generalizability of ML models). 
- Describe the different parts: data budgeting, data pre-processing, model choice and training, assessment of performance.
- Briefly justify and describe your use of simulated data, and results from the pipeline on them.
- Describe results from applying the ML pipeline to the empirical data and what can we learn from them.

Remember: plots are very very important to communicate your process and results.

```{r}

pacman::p_load(tidyverse,purrr, brms, cmdstanr, bayesplot, rstanarm, , viridis, gridExtra, RColorBrewer, reshape2, tidymodels)

set.seed(3)
```

```{r}
color_scheme_set(scheme = "teal")
```

*update_role* 
- assign the role of the column 
update_role(ID, trail, new_role ="second level")

Useful in PCA and when making our recipe




## Part I - Simulating data

Use the meta-analysis reported in Parola et al (2020), *create a simulated dataset* with 100 matched pairs of schizophrenia and controls, each participant producing 10 repeated measures (10 trials with their speech recorded). for each of these "recordings" (data points) produce 10 acoustic measures: 6 from the meta-analysis, 4 with just random noise. Do the same for a baseline dataset including only 10 noise variables. 
Tip: see the slides for the code. 


```{r}
people <- 200
trials <- 10

EffectsMeans_inform <- c(0.25,-0.55,-0.75,-1.26,0.05,1.89,0,0,0,0)
EffectsSD_inform <- c(0.5, 0.29, 0.39,0.63,0.59,0.62,1,1,1,1)

EffectsMeans_noise <- rep(0,10)
EffectsSD_noise <- rep(1,10)

TrialSD <- 0.5
Error <- 0.2
```


```{r}
true_val <- function(means, SDs){
  true_df <- data.frame(Map(rnorm, n=1, mean=means, sd=SDs))
  names(true_df) <- c(paste0("v", seq(10)))
  true_df <- as_tibble(true_df)
}
```


## Simulation of *informed data*

```{r}
#Making a df of individual effect sizes across variables (drawn from "true" means)
sim_d <- data.frame( 
  ID = seq(people/2))

sim_d$values <- map(EffectsMeans_inform, true_val, SDs=EffectsSD_inform) #map effect_means on the true_df as a nested df

sim_d <- unnest(sim_d, cols = values)#Now we have a df with one effect size per variable per participant

```


```{r}
 # Adding trails
trial <- expand.grid(ID=seq(people/2),trial=seq(1,10))

sim_d <- merge(sim_d,trial) 

sim_d <- sim_d %>% 
  rowwise%>% 
  mutate(across(starts_with("v"), ~rnorm(1, mean = rnorm(1,mean=.x, sd=TrialSD), sd = Error), .names="{.col}_obs"))

#Making matched pairs
pairs <- expand.grid(group = c("SZ", "HC"), ID=seq(people/2))

sim_d <- merge(sim_d, pairs)

```


```{r}
#Differentiating between SZ and HC 
matched_data<- sim_d %>% 
  rowwise %>%  
  mutate(across(ends_with("_obs"), ~if_else(group == "SZ", (.x/2), -(.x/2)), .names ="{.col}_matched")) %>%
  select(ID, trial, group, ends_with("_matched"))

```



## Simulation of *noise data*
```{r}
sim_d_noise <- data.frame( 
  ID = seq(people/2))

sim_d_noise$values <- map(EffectsMeans_noise, true_val, SDs=EffectsSD_noise) #map effect_means on the true_df as a nested df

sim_d_noise <- unnest(sim_d_noise, cols = values) #Now we have a df with one effect size per variable per participant
```

```{r}
 # Adding trials
sim_d_noise <- merge(sim_d_noise,trial) 

sim_d_noise <- sim_d_noise %>%
  rowwise %>% 
  mutate(across(starts_with("v"), ~rnorm(1, mean = rnorm(1,mean=.x, sd=TrialSD), sd = Error), .names="{.col}_obs"))

#Making matched pairs
sim_d_noise <- merge(sim_d_noise, pairs)

```


```{r}
#Differentiating between SZ and HC 
matched_noise<- sim_d_noise %>% 
  rowwise %>%  
  mutate(across(ends_with("_obs"), ~if_else(group == "SZ", (.x/2), -(.x/2)), .names ="{.col}_matched")) %>%
  select(ID, trial, group, ends_with("_matched"))

```


#Visualiziing the data :(
```{r}
#Making the df into long format 
long_df <- matched_data %>%
  pivot_longer(cols = ends_with("_matched"), names_to = "variable")

long_df %>% 
  ggplot(aes(x = value, colour=group)) +  
  geom_density() + 
  scale_color_brewer(palette = "Paired") + 
  facet_wrap(~variable) + 
  theme_grey() + 
  ggtitle("Effect sizes of variables")



#Making the df into long format 
long_df_noise <- matched_noise %>%
  pivot_longer(cols = ends_with("_matched"), names_to = "variable")

long_df_noise %>% 
  ggplot(aes(x = value, colour=group)) +  
  geom_density() + 
  scale_color_brewer(palette = "Accent") + 
  facet_wrap(~variable) + 
  theme_grey() + 

  ggtitle("Effect sizes of variables on noise data")

```



## Part II - ML pipeline on simulated data

On the two simulated datasets (separately) build a machine learning pipeline: 
*i)* create a data budget (e.g. balanced training and test sets) 

*ii)* pre-process the data (e.g. scaling the features)

*iii)* fit and assess a classification algorithm on the training data (e.g. Bayesian multilevel logistic regression)

*iv)* assess performance on the test set

*v)* discuss whether performance is as expected and feature importance is as expected.


Bonus question: replace the bayesian multilevel regression with a different algorithm, e.g. SVM or random forest (but really, anything you'd like to try).



#Data budgetting and preprocessing
```{r}
#Splitting data 80/20
spl <- sample(seq(people/2), 20)

train_inf <- matched_data %>% subset(!(ID %in% spl))
test_inf <- matched_data %>% subset(ID %in% spl)

```


```{r}
#Centering the training data
rec_informed <- train_inf %>% recipe(group ~.) %>% 
  update_role(ID, trials, new_role = "Second level") %>%
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) %>%
  prep(train=train_inf, retain = T)
  

#Using the recipe on the test data
bake(rec_informed, new_data = test_inf)
```





#Choosing a classification algorithm - Logistic Regression
```{r}
#Setting up the formula(s)

f1 <- bf(group ~ 1 + v1_obs_matched+ v2_obs_matched + v3_obs_matched + v4_obs_matched + v5_obs_matched + v6_obs_matched + v7_obs_matched + v8_obs_matched + v9_obs_matched + v10_obs_matched + (1+trial|ID))
#Group is a function of each of the variables

variables <- train_inf %>% select(c(starts_with("v"))
                                  
f1 <- bf(group ~ 1 + variables+ (1+trial|ID))


```

```{r}

get_prior(data=train_inf, 
          family = bernoulli, 
          f1)

p1 <- c(
  prior(normal(0,1), class=Intercept),
  prior(normal(0,0.3), class=b), 
  prior(normal(0,0.3), class= sd)
)
```

```{r}

m1_prior <-brm(
  f1,
  data=train_inf,
  family = bernoulli(link = "logit"),
  prior = p1,
  sample_prior = "only",
  #backend = "cmdstanr",
  #threads = threading(2),
  iter = 2000,
  chains = 2,
  cores = 2,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)



pp_check(m1_prior, ndraws =1000)
```

```{r}
m1 <- brm(
  f1,
  data=train_inf,
  family = bernoulli,
  prior = p1,
  sample_prior = T,
  #backend = "cmdstanr",
  #threads = threading(2),
  iter = 2000,
  chains = 2,
  cores = 2,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)

```





## Part III - Applying the ML pipeline to empirical data

Download the empirical dataset from brightspace and apply your ML pipeline to the new data, adjusting where needed. *Warning*: in the simulated dataset we only had 10 features, now you have many more! Such is the life of the ML practitioner. Consider the impact a higher number of features will have on your ML inference, and decide whether you need to cut down the number of features before running the pipeline (or alternatively expand the pipeline to add feature selection).

Data: https://www.dropbox.com/s/7ky1axvea33lgye/Ass3_empiricalData1.csv?dl=0
